---
title: "Statistical Inference - Course project Part 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Inference - Course project Part 1

######(from assignment)
In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. I will do a thousand simulations.

##Exploration
What's a rexp? - let's do 1000 values and plot them
```{r, warning=F}
library(ggplot2)
set.seed(10101)

#Since this rexp thing is probably random, I will save it
myDist <- rexp(1000,.2)
head(myDist)
summary(myDist)
#There were warnings asking to set binwidth, so I did it.
qplot(myDist, binwidth=1)+geom_vline(xintercept = median(myDist)) +
  ggtitle('Just a rexp example for 1000 counts')

```

##Simulations
Now let's do 1000 of these with size of 40 per project requirement.
I will save them to a matrix with 1000 rows and 40 columns
```{r}
samples <-1000
n <- 40
lambda <- 0.2
dataSet <-matrix(data=rexp(n * samples, lambda), nrow=samples)
```

Let's take means of every row (observation)
```{r}
means <- apply(dataSet, 1, mean)
summary(means)

#Let's plot them too!
g<-qplot(means, binwidth=.1)+geom_vline(xintercept = median(means))
g
```

Mean of means turns out to be different every time I run knitr - I should set seed up top.
With seed 10101 mean of means is constant - 4.966

##Sample Mean versus Theoretical Mean
Condition of the project is that mean is 1/lambda
```{r}
1/lambda
```

**Theoretical mean 5 and sampled mean 4.966 are pretty close!**

##Sample Variance versus Theoretical Variance
Let's find theoretical and practical standard deviation and  variance
```{r}
theo_sd <- (1/lambda)/sqrt(n)
theo_sd

theo_var <- theo_sd^2
theo_var

prac_sd <- sd(means)
prac_sd
prac_var <- var(means)
prac_var
```

Here we have sd off by couple of decimals(0.7681136 vs 0.7905694), and variance of course in the same area. I wish I had a few years of statistical background to conclude it it close enough, but I also have a gut feeling that in scope of this project the practice will fit the theory, so they are close.

##Distribution
Here's a overlay with a random distribution with mean = 4.996 and sd = 0.7905694.
It fits really well, so we conclude the distribution of means of rexp is approximately normal.

```{r, warning=F,message=FALSE}
line <- rnorm(1000,5,0.7905694)
df <- data.frame(means=means,line=line)
ggplot(df, aes(means)) + 
  geom_histogram(aes(means, bindwidth=.1, fill = "black",alpha = 0.2)) + 
  geom_histogram(aes(line, bindwidth=.1, fill = "red", alpha = 0.2)) +
  ggtitle('Overlay of black exponential means with red normal distribution')
```




Someone else's graph - beautiful overlay
# density
y <- dnorm(x = x, mean = mean(x), sd = sd(x))
x <- as.data.frame(x)
m <- ggplot(x, aes(x=x))
# sample histogram bars
m <- m + geom_histogram(binwidth = 0.3, aes(y = ..density..), col="blue", fill="blue",
alpha = 0.4)
# normal curve
m <- m + geom_area(aes(y = y), alpha = 0.3, fill="yellow", col="navy", size = 0.8)
# visual comparison of the means
m <- m + geom_vline(xintercept = mean(x$x), col="darkblue", size = 3.5)
m <- m + geom_vline(xintercept = 1/lambda, col="orange", size = 1.5, linetype = "dashed")
# labels
m <- m + ggtitle(label = "Sample vs. Theoretical Distribution")
m <- m + labs(x = "Sample Mean", y = "Density")
m <- m + scale_x_continuous(breaks = c(0:round(sample_mean + (sqrt(sample_var) * 3))))
# render
m
